\documentclass[czech,master]{diploma}

% Standard packages
\usepackage[autostyle=true,czech=quotes]{csquotes} % correct quotation typesetting, support for biblatex
\usepackage[backend=biber, style=iso-numeric, alldates=iso]{biblatex} % bibliography
\usepackage{dcolumn} % table columns with numeric values
\usepackage{subfig} % macros for "subfigures" and "subtables"

% Custom packages
\usepackage{float} % better figure placement (H)
\usepackage{bookmark} % Required for generating PDF bookmarks
\usepackage{hyperref} % Links in PDF
\usepackage{subcaption} % For multiple images in one float
\usepackage{tikz} % For drawing in LaTeX
\usepackage{amsmath} % For improved equation typesetting
\usepackage{tablefootnote} % For footnotes in tables
\usepackage{color} % For colored text (e.g., TO.DO notes)
\usepackage{array} % For advanced table formatting
\usepackage{amsfonts} % For additional math symbols
\usepackage[cpp]{diplomalst}
\usepackage{fontspec}
\usepackage{pgfmath}
\usepackage{additionallst}
\usepackage{minted}

% ==] Overrides and custom settings [==

% Enable fixed-point arithmetic in TikZ for more accurate calculations (e.g., for cost computations)
\usetikzlibrary{fpu}

% Table of contents depth - TODO: Remove in final version!
\setcounter{tocdepth}{3}

% New type of table column with numbers aligned by decimal comma
\newcolumntype{d}[1]{D{,}{,}{#1}}

% Settings for minted code listings
\setminted{fontsize=\fontsize{9}{11}\selectfont, baselinestretch=1, frame=lines, framesep=8pt, linenos, breaklines, ignorelexererrors=true}
\renewcommand\listingscaption{Výpis}
\renewcommand{\lstlistingname}{Prompt}

% ==] Thesis metadata and configuration [==

\ThesisAuthor{Bc. Pavel Mikula}
\ThesisSupervisor{Mgr. Ing. Michal Krumnikl, Ph.D.}
\CzechThesisTitle{Integrace velkých jazykových modelů do systému Kelvin}
\EnglishThesisTitle{Integrating Large Language Models into Kelvin}
\SubmissionYear{2026}

\ThesisAssignmentFileName{resources/ThesisSpecification_MIK0486.pdf}

\CzechAbstract{
    Tato diplomová práce se zabývá návrhem a implementací modulu pro automatickou analýzu zdrojových kódů pomocí velkých jazykových modelů (LLM) v rámci webového informačního systému využívaného při výuce programování.
    Práce popisuje principy fungování LLM, architekturu Transformer a základní metody jejich využití při analýze a hodnocení kódu.
    Součástí je porovnání různých způsobů nasazení a přístupů k promptování z hlediska kvality výsledků, provozních nároků a ochrany dat.
    Navržené řešení je experimentálně ověřeno na reálných datech a jeho výsledky ukazují možnosti využití LLM jako podpůrného nástroje při hodnocení programátorských úloh.
}
\CzechKeywords{velké jazykové modely, analýza zdrojového kódu, automatické hodnocení, Transformer, promptování, výuka programování}

\EnglishAbstract{
    This thesis focuses on the design and implementation of a module for automatic analysis of source code using large language models (LLMs) within a web-based information system used in programming education.
    The thesis describes the principles of LLMs, the Transformer architecture, and approaches for applying these models to source code analysis.
    It also compares different deployment approaches and prompting strategies with respect to output quality, operational requirements, and data protection.
    The proposed solution is experimentally evaluated on real-world data, demonstrating the potential of LLMs as a supporting tool for assessing programming assignments.
}
\EnglishKeywords{large language models, source code analysis, automated assessment, Transformer, prompting, programming education}

% TODO: Check if all these acronyms are actually used in the text, and remove those that are not.
\AddAcronym{LLM}{Large Language Model}
\AddAcronym{API}{Application Programming Interface}
\AddAcronym{AST}{Abstract Syntax Tree}
\AddAcronym{BPE}{Byte Pair Encoding}
\AddAcronym{CD}{Continuous Deployment}
\AddAcronym{CFG}{Control Flow Graph}
\AddAcronym{CI}{Continuous Integration}
\AddAcronym{CoT}{Chain-of-Thought}
\AddAcronym{CPU}{Central Processing Unit}
\AddAcronym{CUDA}{Compute Unified Device Architecture}
\AddAcronym{DPA}{Data Processing Agreement}
\AddAcronym{GDPR}{General Data Protection Regulation}
\AddAcronym{GGUF}{GPT-Generated Unified Format}
\AddAcronym{GPU}{Graphics Processing Unit}
\AddAcronym{GPT}{Generative Pre-trained Transformer}
\AddAcronym{JSON}{JavaScript Object Notation}
\AddAcronym{KL}{Kullback-Leibler}
\AddAcronym{LSTM}{Long Short-Term Memory}
\AddAcronym{PPO}{Proximal Policy Optimization}
\AddAcronym{PR}{Pull Request}
\AddAcronym{REST}{Representational State Transfer}
\AddAcronym{RLHF}{Reinforcement Learning from Human Feedback}
\AddAcronym{RNN}{Recurrent Neural Network}
\AddAcronym{SFT}{Supervised Fine-Tuning}
\AddAcronym{SLA}{Service Level Agreement}
\AddAcronym{UI}{User Interface}
\AddAcronym{UX}{User Experience}
\AddAcronym{VRAM}{Video Random Access Memory}
\AddAcronym{DOLOS}{Detection of Logical Similarity}
\AddAcronym{MOSS}{Measure Of Software Similarity}

\Acknowledgement{Rád bych na tomto místě poděkoval svému vedoucímu práce Mgr. Ing. Michalu Krumniklovi, Ph.D. za odborné vedení, cenné rady a trpělivost při konzultacích v průběhu celé práce. Dále děkuji kolegům, kteří se podíleli na hodnocení výstupů jazykových modelů a svými připomínkami přispěli k získání objektivnějších výsledků experimentálního ověření.}

\addbibresource{resources/sauce.bib}

% ==] Custom macros [==

\input{macros}

%
% ==] Main document [==
%

\begin{document}
    \MakeTitlePages

    % Pictures
    \listoffigures
    \clearpage

    % Tables
    \listoftables
    \clearpage

    % Source codes
    \renewcommand{\lstlistlistingname}{Seznam výpisů}
    \addcontentsline{toc}{chapter}{Seznam výpisů}
    \lstlistoflistings

    % Introduction
    \clearpage
    \input{chapters/1-Introduction}

    % Main chapters
    \input{chapters/2-Kelvin}
    \input{chapters/3-LLM}
    \input{chapters/4-Selection}
    \input{chapters/5-Implementation}
    \input{chapters/6-Summary}
    \input{chapters/7-Suggestion}
    \input{chapters/8-Conclusion}

    \appendix
    \input{chapters/9-Appendix}

    % Bibliography
    \printbibliography[title={Literatura}, heading=bibintoc]

\end{document}
